Quantum Hoare Type Theory

### 2020-06-03 ms-paper/ms-paper.tex
Changes to everything apart from the main paper

@@

This paper is organized as follows. We discuss background in the next section (\cref{sec:bg}); specifically, quantum computing (\cref{sec:qc}),  Hoare Type Theory in the classical setting (\cref{sec:htt}) and then the Quantum IO Monad interface (\cref{sec:qio}). Then we introduce our contributions in merging these ideas together for specification and enforcement of useful semantic properties in section \cref{sec:qhtt} along with its syntax (\cref{sec:syntax}), typing (\cref{sec:typing}) and, some examples (\cref{sec:examples}). Some related work follows in \cref{related}. Finally, we conclude and share ideas for future work in the final section (\cref{sec:conclusion}).

@@

As may be apparent, initialization and measurement are analogous to HTT's \texttt{alloc} and \texttt{dealloc} primitives. We are working out precise details for unitary application, but at a high-level, they involve ensuring that the given unitary term actually represents a unitary matrix and updating the quantum state with the result of unitary application. This involves incorporating the monoidal structure of unitaries into our theory (as is done in QIO). This step also involves generating verification conditions that cannot be checked during this first phase of typechecking.

@@

\section{Related Work}
\label{related}
%\textit{Objective of this section\ldots}

@@

\begin{acks}
%	I would like to thank my advisor John Reppy for suggesting me to look at Hoare Type Theory that led to this master's thesis and for his constant support and encouragement to see it through. I also thank my committee members Fred Chong and Ravi Chugh for reading drafts of this document and providing useful feedback.

>>>>

% modified from bidir-survey
\definecolor{dRed}{rgb}{0.85, 0.0, 0.0}
\definecolor{dBlue}{rgb}{0.0, 0.0, 0.85}

\newcommand{\chkcolor}{dBlue}
\newcommand{\syncolor}{dRed}
\newcommand{\chk}{\,\textcolor{\chkcolor}{\Leftarrow}\,}
\newcommand{\uncoloredsyn}{{\Rightarrow}}
\newcommand{\syn}{\,\textcolor{\syncolor}{\uncoloredsyn}\,}

@@

This paper is organized as follows. We discuss background in the next section (\cref{sec:bg}); specifically, quantum computing (\cref{sec:qc}),  Hoare Type Theory in the classical setting (\cref{sec:htt}) and then the Quantum IO Monad interface (\cref{sec:qio}). Then we introduce our contributions in merging these ideas together for specification and enforcement of useful semantic properties in section \cref{sec:qhtt} along with its syntax (\cref{sec:syntax}), typing (\cref{sec:typing}) and, some examples (\cref{sec:examples}). Some related work follows in \cref{sec:related}. Finally, we conclude and share ideas for future work in the final section (\cref{sec:conclusion}).

@@

As may be apparent, initialization and measurement are analogous to HTT's \texttt{alloc} and \texttt{dealloc} primitives. We are working out precise details for unitary application, but at a high-level, they involve ensuring that the given unitary term actually represents a unitary matrix and updating the quantum state with the result of unitary application. This involves incorporating the monoidal structure of unitaries into our theory (as is done in QIO). This step also involves generating verification conditions that cannot be checked during this first phase of typechecking. We are also looking at alternatives such as the recently proposed sum-over-paths action semantics \cite{amy18}.

@@

\section{Related Work}
\label{sec:related}
%\textit{Objective of this section\ldots}

@@

\begin{acks}
	I thank John Reppy, Robert Rand and the anonymous reviewers of QPL 2020 for their feedback on an earlier presentation of this work.
